---
title: "README"
output: html_document
---
```{r tablePrep}
#| include = FALSE
library(tidyverse)
library(here)
library(knitr)
library(kableExtra)

# 1. Read the CSV file using here()
data <- read_csv(here("data/capex987_20260128/jasco_fieldDefinitions.csv"))
```

# PRELIMINARY datasets for Slocum Glider with JASCO OceanObserver  'capex987'

Non-audio data from Slocum Glider with JASCO OceanObserver was prepared using [CHECK] the Post-mission processing via SWFSC Ecosystem Science Division's [Slocum Glider Data Processing Methods)](https://swfsc.github.io/glider-lab-manual/content/slocum-glider-data.html). Below are descriptions of the available datasets:

## capex987_20260128_flight_timeseries_engineering.csv

The flight engineering dataset contains information on the pitch, roll, heading, etc. for each dive. Methods to develop this dataset can be found here (cite!), and definitions/units for data are found in the table below. *Note: Data are preliminary and require additional QAQC.*


```{r engineering}
#| echo = FALSE
#| 
# 2. Render and style the table
data %>% 
  filter(Dataset == "flight_timeseries_engineering") %>%
  kbl(
    align = "l", # Justify text to the left (use "c" for center or "r" for right)
    booktabs = TRUE # Clean, professional tablines
  ) %>%
  kable_styling(
    latex_options = c("striped", "HOLD_position"), # Alternating row shading (gray/white)
    bootstrap_options = c("striped", "hover", "condensed"), # Web alternative if compiling to HTML
    full_width = TRUE # Fits text across the full page width with automatic word wrap
  ) %>%
  row_spec(
    0, # Target header row
    bold = TRUE,
    extra_css = "border-bottom: 3px solid black;" # Thick bottom border for HTML
  ) %>%
  # Apply thick line underneath header for PDF/LaTeX output
  {
    if (knitr::is_latex_output()) {
      row_spec(., 0, bold = TRUE) %>%
        row_spec(0, extra_latex_after = "\\hline\\noalign{\\hrule height 1.5pt}")
    } else {
      .
    }
  }
```

 
## capex987_20260128_GPS_timeseries.csv

GPS time series data is formatted with one row per dive, with start and end times and locations for each dive. Movement between the end of one dive and the start of the next is due to any surface drift. The GPS dataset contains information on the Start and End Time, Latitude, Longitude as well as the duration, distance, and max depth for each dive. Additional variables are presented in the table (below). Methods to develop this dataset can be found here (cite!), and definitions/units for data are found in the table below. *Note: Data are preliminary and require additional QAQC.*

```{r gps}
#| echo = FALSE
#| 
data %>% 
  filter(Dataset == "GPS_timeseries") %>%
  kbl(
    align = "l", # Justify text to the left (use "c" for center or "r" for right)
    booktabs = TRUE # Clean, professional tablines
  ) %>%
  kable_styling(
    latex_options = c("striped", "HOLD_position"), # Alternating row shading (gray/white)
    bootstrap_options = c("striped", "hover", "condensed"), # Web alternative if compiling to HTML
    full_width = TRUE # Fits text across the full page width with automatic word wrap
  ) %>%
  row_spec(
    0, # Target header row
    bold = TRUE,
    extra_css = "border-bottom: 3px solid black;" # Thick bottom border for HTML
  ) %>%
  # Apply thick line underneath header for PDF/LaTeX output
  {
    if (knitr::is_latex_output()) {
      row_spec(., 0, bold = TRUE) %>%
        row_spec(0, extra_latex_after = "\\hline\\noalign{\\hrule height 1.5pt}")
    } else {
      .
    }
  }
```



## capex987_20260128_science_timeseries.csv

The science dataset contains information on the depth, temperature, salinity, density, and sound velocity for each dive. Methods to develop this dataset can be found here (cite!), and definitions/units for data are found in the table below. *Note: Data are preliminary and require additional QAQC.*

```{r sci}
#| echo = FALSE
#| 
data %>% 
  filter(Dataset == "science_timeseries") %>%
  kbl(
    align = "l", # Justify text to the left (use "c" for center or "r" for right)
    booktabs = TRUE # Clean, professional tablines
  ) %>%
  kable_styling(
    latex_options = c("striped", "HOLD_position"), # Alternating row shading (gray/white)
    bootstrap_options = c("striped", "hover", "condensed"), # Web alternative if compiling to HTML
    full_width = TRUE # Fits text across the full page width with automatic word wrap
  ) %>%
  row_spec(
    0, # Target header row
    bold = TRUE,
    extra_css = "border-bottom: 3px solid black;" # Thick bottom border for HTML
  ) %>%
  # Apply thick line underneath header for PDF/LaTeX output
  {
    if (knitr::is_latex_output()) {
      row_spec(., 0, bold = TRUE) %>%
        row_spec(0, extra_latex_after = "\\hline\\noalign{\\hrule height 1.5pt}")
    } else {
      .
    }
  }
```

## capex987_20260128_profiles.csv

Profile data.


```{r profile}
#| echo = FALSE
#| 
data %>% 
  filter(Dataset == "profile") %>%
  kbl(
    align = "l", # Justify text to the left (use "c" for center or "r" for right)
    booktabs = TRUE # Clean, professional tablines
  ) %>%
  kable_styling(
    latex_options = c("striped", "HOLD_position"), # Alternating row shading (gray/white)
    bootstrap_options = c("striped", "hover", "condensed"), # Web alternative if compiling to HTML
    full_width = TRUE # Fits text across the full page width with automatic word wrap
  ) %>%
  row_spec(
    0, # Target header row
    bold = TRUE,
    extra_css = "border-bottom: 3px solid black;" # Thick bottom border for HTML
  ) %>%
  # Apply thick line underneath header for PDF/LaTeX output
  {
    if (knitr::is_latex_output()) {
      row_spec(., 0, bold = TRUE) %>%
        row_spec(0, extra_latex_after = "\\hline\\noalign{\\hrule height 1.5pt}")
    } else {
      .
    }
  }
```


## capex987_20260128_FakeData_EventDetections.csv
Due to complications beyond our control, accurate acoustic detection of cetacean species via manual and/or automated methods could not be completed prior to the Hackathon. A set of synthesized (fake!) data was provided in the data/detection folder to provide sample formatted data if a group would like to lead a project working with marine mammal acoustic data (projects developed with these synthesized datasets should be reproducible with real datasets in the future!). Field definitions/units for data are found in the table below (fields are the same for EventDetections and HrlyDetections).

Event Detections (ex, sg607_20260128_FakeData_HrlyDetections.csv) These files show different events with a start and end date/time that may be greater than or less than an hour. So, some long humpback whale events in this sheet may extend through several hourly bins from the Hourly Detection file.

## capex987_20260128_FakeData_HrlyDetections.csv
Hourly Detections (ex, FakeData_HrlyDetections_GliderRodeo2026_Belladonna.csv) These files show positive detections in hourly bins. Effort with no detections are not provided.

```{r det}
#| echo = FALSE
#| 
data %>% 
  filter(Dataset == "detections") %>%
  kbl(
    align = "l", # Justify text to the left (use "c" for center or "r" for right)
    booktabs = TRUE # Clean, professional tablines
  ) %>%
  kable_styling(
    latex_options = c("striped", "HOLD_position"), # Alternating row shading (gray/white)
    bootstrap_options = c("striped", "hover", "condensed"), # Web alternative if compiling to HTML
    full_width = TRUE # Fits text across the full page width with automatic word wrap
  ) %>%
  row_spec(
    0, # Target header row
    bold = TRUE,
    extra_css = "border-bottom: 3px solid black;" # Thick bottom border for HTML
  ) %>%
  # Apply thick line underneath header for PDF/LaTeX output
  {
    if (knitr::is_latex_output()) {
      row_spec(., 0, bold = TRUE) %>%
        row_spec(0, extra_latex_after = "\\hline\\noalign{\\hrule height 1.5pt}")
    } else {
      .
    }
  }
```





