---
title: "Field Definitions Completeness Report"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    theme: flatly
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

library(tidyverse)
library(here)
library(DT)

# Pins the project root to the folder containing this .Rmd file, regardless
# of the working directory in effect when knitting (this is what avoids the
# "cannot open the connection" error from here() guessing the wrong root).
here::i_am("field_definitions_report.Rmd")

source(here::here("R", "check_field_definitions.R"))
```

This report checks whether every column in your datasets is documented in
your field definitions file, and whether every documented field has a
complete `Definition` and `Units` value.

## Inputs {.tabset}

Edit the chunk below to point to your own datasets and field definitions
file. Paths are relative to the project root (resolved via `here::here()`).

```{r inputs, echo=TRUE}
# ------------------------- EDIT THIS CHUNK -------------------------------
datasets <- list(
  detections = here("../data/capex987_20260128/capex987_20260128_FakeData_EventDetections.csv"),
  detections = here("../data/capex987_20260128/capex987_20260128_FakeData_HrlyDetections.csv"),
  # flight_timeseries_engineering = here("../data/risso_20260128/risso-20260128_flight_timeseries_engineering.csv"),
  # flight_timeseries_modeled = here("../data/stenella_20260128/risso-20260128_flight_timeseries_modeled.csv"),
  profiles = here("../data/capex987_20260128/capex987-20260128-delayed-profiles.csv"),
  GPS_timeseries = here("../data/capex987_20260128/capex987-20260128_GPS_timeseries.csv")
  # science_timeseries = here("../data/capex987_20260128/capex987_20260128_science_timeseries.csv")
)

field_definitions_path <- here("../data/capex987_20260128/slocum_fieldDefinitions.csv")
# ---------------------------------------------------------------------------
```

```{r run}
results <- check_field_definitions(datasets, field_definitions_path)
```

## Summary

```{r summary-table}
summary_tbl <- tibble::tibble(
  Check = c(
    "Columns in data, undocumented in fieldDefinitions",
    "Columns in fieldDefinitions, not found in data",
    "Dataset names in fieldDefinitions not supplied to this report",
    "Dataset names supplied but absent from fieldDefinitions",
    "fieldDefinitions rows missing Definition and/or Units"
  ),
  Count = c(
    nrow(results$missing_from_defs),
    nrow(results$missing_from_data),
    length(results$datasets_only_in_defs),
    length(results$datasets_only_in_data),
    nrow(results$missing_info)
  )
)

DT::datatable(summary_tbl, options = list(dom = "t"), rownames = FALSE)
```

```{r summary-flag, results='asis'}
if (all(summary_tbl$Count == 0)) {
  cat("**All checks passed — every field is documented and complete.**")
} else {
  cat("**", sum(summary_tbl$Count), "issue(s) found. See details below.**")
}
```

## 1. Columns Missing From fieldDefinitions

Columns that exist in a dataset but have no matching row in the field
definitions file.

```{r missing-from-defs, results='asis'}
if (nrow(results$missing_from_defs) == 0) {
  cat("None — every dataset column is documented.")
} else {
  DT::datatable(results$missing_from_defs, rownames = FALSE)
}
```

## 2. Columns Missing From the Data

Rows in the field definitions file that reference a Dataset / Column_Name
combination that was not found in the corresponding dataset (either the
column doesn't exist, or is misspelled).

```{r missing-from-data, results='asis'}
if (nrow(results$missing_from_data) == 0) {
  cat("None — every documented field was found in the data.")
} else {
  DT::datatable(results$missing_from_data, rownames = FALSE)
}
```

## 3. Dataset Name Mismatches

Checks that the names used in `datasets` line up with the values in the
`Dataset` column of the field definitions file.

```{r dataset-mismatch, results='asis'}
if (length(results$datasets_only_in_defs) == 0 && length(results$datasets_only_in_data) == 0) {
  cat("None — dataset names match between your input list and fieldDefinitions.")
} else {
  if (length(results$datasets_only_in_defs) > 0) {
    cat(
      "**In fieldDefinitions, but not supplied to this report:** ",
      paste(results$datasets_only_in_defs, collapse = ", "),
      "\n\n"
    )
  }
  if (length(results$datasets_only_in_data) > 0) {
    cat(
      "**Supplied to this report, but absent from fieldDefinitions:** ",
      paste(results$datasets_only_in_data, collapse = ", ")
    )
  }
}
```

## 4. Missing Definition / Units

Field definition rows where `Definition` and/or `Units` is blank.

```{r missing-info, results='asis'}
if (nrow(results$missing_info) == 0) {
  cat("None — every documented field has both a Definition and a Units value.")
} else {
  results$missing_info |>
    dplyr::mutate(
      Missing = dplyr::case_when(
        Definition_missing & Units_missing ~ "Definition & Units",
        Definition_missing ~ "Definition",
        Units_missing ~ "Units"
      )
    ) |>
    dplyr::select(Dataset, Column_Name, Missing) |>
    DT::datatable(rownames = FALSE)
}
```
