---
title: "README"
output: html_document
---

**Glider Rodeo Synthesized (fake!) Marine Mammal Detection Data**

Due to complications beyond our control, accurate acoustic detection of cetacean species via manual and/or automated methods could not be completed prior to the Hackathon. A set of synthesized (fake!) data was provided in the data/detection folder to provide sample formatted data if a group would like to lead a project working with marine mammal acoustic data (projects developed with these synthesized datasets should be reproducible with real datasets in the future!)

**Datasets:**

1.  Hourly Detections (ex, FakeData_HrlyDetections_GliderRodeo2026_Belladonna.csv) These files show positive detections in hourly bins. Effort with no detections are not provided.

2.  Event Detections (ex, FakeData_EventDetections_GliderRodeo2026_Belladonna.csv) These files show different events with a start and end date/time that may be greater than or less than an hour. So, some long humpback whale events in this sheet may extend through several hourly bins from the Hourly Detection file.

**Dataset Field Definitions:**

Datasets are formatted for NOAA's Makara database. Not all fields are provided in our synthesized datasets. If your Hackathon project requires information from one of these fields, you can add these fields (using these specific formats/definitions) using code (please do not modify the existing datasets provided!). Below are field descriptions and formatting:

```{r tablePrep}
#| include = FALSE
library(tidyverse)
library(here)
library(knitr)
library(kableExtra)

# 1. Read the CSV file using here()
data <- read_csv(here("data/detection/SWFSC_YYYYMMDD_DETECTIONDATA_FieldDefinitions.csv"))
```

```{r fieldNames}
#| echo = FALSE
#| 
# 2. Render and style the table
data %>%
  kbl(
    align = "l", # Justify text to the left (use "c" for center or "r" for right)
    booktabs = TRUE # Clean, professional tablines
  ) %>%
  kable_styling(
    latex_options = c("striped", "HOLD_position"), # Alternating row shading (gray/white)
    bootstrap_options = c("striped", "hover", "condensed"), # Web alternative if compiling to HTML
    full_width = TRUE # Fits text across the full page width with automatic word wrap
  ) %>%
  row_spec(
    0, # Target header row
    bold = TRUE,
    extra_css = "border-bottom: 3px solid black;" # Thick bottom border for HTML
  ) %>%
  # Apply thick line underneath header for PDF/LaTeX output
  {
    if (knitr::is_latex_output()) {
      row_spec(., 0, bold = TRUE) %>%
        row_spec(0, extra_latex_after = "\\hline\\noalign{\\hrule height 1.5pt}")
    } else {
      .
    }
  }
```
